---
title: "Police stop and search data extraction"
author: "Black Thrive Global Research Team"
date: "January 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview 

This script retrieves stop and search data for a specified set of Local Authority Districts (LADs) in Britain over a specified time period. This is achieved in several steps.

1. Coordinates defining the geographic boundary of each LAD are extracted from a datafile obtained from the Office for National Statistics and formatted in a way that can be used by an extraction function to obtain police data
2. A custom functions reads the coordinates and queries the Police API for stop and search records within each LAD over the specified time period.
3. Data on which Police Force serves each LAD are added to the data acquired from the function

# Packages

```{r}
packages <- c('tidyverse',
              'jsonify', # for reading json
              'leaflet', # for map
              'httr', # for http requests
              'data.table', # for unlisting
              'rgdal', # for geopackage
              'sf', # for geopackage
              'chron' # for dealing with time values
)
pkg_notinstall <- packages[!(packages %in% installed.packages()[,"Package"])]
lapply(pkg_notinstall, install.packages, dependencies = TRUE)
lapply(packages, library, character.only = TRUE)
```

# Specify coordinates

This section of the script specifies the coordinates defining the geographic boundary of each LAD and formats them into a list that can be used by the data extraction function. It also adds county, region, and country data to the coordinates where possible.


## Read data

```{r}
# read boundary data
coords <- geojsonR::FROM_GeoJson("../data/Local_Authority_Districts_(December_2021)_UK_BFE.geojson")

# read list of local authorities
las <- read_csv("../data/ward_la_county_region_country_2021.csv")

# just select by la and get rid of wards 
las_no_ward <- las[!duplicated(las$LAD21NM),] %>%
  select(-c(1:2))

las <- las_no_ward

# list of which forces serve which LADs
forces_df <- read_csv("../data/la_to_force_2020.csv")
# just select by la and get rid of CSPs 
forces_df <- forces_df[!duplicated(forces_df$LAD20NM),] %>%
  select(-c(4:5))

```

```{r}
# check that NI LAs are in this coord set

# targets <- c("Causeway Coast and Glens",
#              "Derry City and Strabane",
#              "Fermanagh and Omagh",
#              "Mid Ulster",
#              "Mid and East Antrim",
#              "Antrim and Newtownabbey",
#              "Armagh City, Banbridge and Craigavon",
#              "Lisburn and Castlereagh",
#              "Belfast",
#              "Ards and North Down",
#              "Newry, Mourne and Down")
# count <- 0
# for(i in 1:374){
#   if(!is.na(match(coords[["features"]][[i]][["properties"]][["LAD21NM"]], targets))){
#     count <- count + 1
#   }
# }
# count
```

## Format coordinates

Here the coordinate list is structured into the appropriate format for the extraction function, adding data on county, region, and country. The script also records missing elements of data (e.g., missing counties, regions, etc.) for manual inspection.

```{r}
# intialise vector for extracted coordinates
coords_extracted <- vector(mode = "list", length = length(coords[["features"]]))

# intialise data frame to record missing data
missing_data_df <- data.frame(setNames(rep(list(NA), 5), c("Index","Name","County_missing","Region_missing","Country_missing")))

for(i in 1:length(coords[["features"]])){
  coords_extracted[[i]][["la_name"]] <- coords[["features"]][[i]][["properties"]][["LAD21NM"]]
  coords_extracted[[i]][["la_code"]] <- coords[["features"]][[i]][["properties"]][["LAD21CD"]]
  #coords_extracted[[i]][["coords"]] <- lapply(coords_extracted[[i]][["coords"]][[1]], unlist)
  
  # add la, county, region, and country names
  try(
    coords_extracted[[i]][["county"]] <- as.character(las[which(las$LAD21CD == coords_extracted[[i]][["la_code"]]), "CTY21NM"])
  )
  try(
    coords_extracted[[i]][["region"]] <- as.character(las[which(las$LAD21CD == coords_extracted[[i]][["la_code"]]), "RGN21NM"])
  )
  try(
    coords_extracted[[i]][["country"]] <- as.character(las[which(las$LAD21CD == coords_extracted[[i]][["la_code"]]), "CTRY21NM"])
  )
  try(
    coords_extracted[[i]][["force"]] <- as.character(forces_df[which(forces_df[,"LAD20CD"] == coords_extracted[[i]][["la_code"]]), "PFA20NM"])
  )
  
  if(coords_extracted[[i]][["force"]] == "character(0)"){
    coords_extracted[[i]][["force"]] <- NA
  }
  names(coords_extracted)[i] <- coords_extracted[[i]][["la_name"]]
  
  coords_extracted[[i]][["coords"]] <- coords[["features"]][[i]][["geometry"]][["coordinates"]]

  # report which LAs have missing values 
  if(is.na(coords_extracted[[i]][["county"]])){
    #print(paste0(names(coords_extracted)[i], " county missing (", i, ")"))
    missing_data_df[i, "Index"] <- i
    missing_data_df[i, "Name"] <- coords_extracted[[i]][["la_name"]]
    missing_data_df[i, "Country"] <- coords_extracted[[i]][["country"]]
    missing_data_df[i, "County_missing"] <- 1
  }
  if(is.na(coords_extracted[[i]][["region"]])){
    #print(paste0(names(coords_extracted)[i], " region missing (", i, ")"))
    missing_data_df[i, "Index"] <- i
    missing_data_df[i, "Name"] <- coords_extracted[[i]][["la_name"]]
    missing_data_df[i, "Country"] <- coords_extracted[[i]][["country"]]
    missing_data_df[i, "Region_missing"] <- 1
  }
  if(is.na(coords_extracted[[i]][["country"]])){
    #print(paste0(names(coords_extracted)[i], " country missing (", i, ")"))
    missing_data_df[i, "Index"] <- i
    missing_data_df[i, "Name"] <- coords_extracted[[i]][["la_name"]]
    missing_data_df[i, "Country"] <- coords_extracted[[i]][["country"]]
    missing_data_df[i, "Country_missing"] <- 1
  }
  if(is.na(coords_extracted[[i]][["force"]])){
    #print(paste0(names(coords_extracted)[i], " country missing (", i, ")"))
    missing_data_df[i, "Index"] <- i
    missing_data_df[i, "Name"] <- coords_extracted[[i]][["la_name"]]
    missing_data_df[i, "Country"] <- coords_extracted[[i]][["country"]]
    missing_data_df[i, "Force_missing"] <- 1
  }

  
  for(j in 1:length(coords_extracted[[i]][["coords"]])){
    if(is.list(coords_extracted[[i]][["coords"]][[j]])){
      coords_extracted[[i]][["coords"]] <- lapply(coords_extracted[[i]][["coords"]][[1]], unlist)
    }
    
  
  #coords_extracted[[i]][["coords"]][[j]] <- as.data.frame(coords_extracted[[i]][["coords"]][[j]])
  #colnames(coords_extracted[[i]][["coords"]][[j]]) <- c("long","lat")
  }
  
  
}

# drop rows from missing_data_df that are NA
missing_data_df <- missing_data_df[rowSums(is.na(missing_data_df)) != ncol(missing_data_df),]

# unlist coordinates to make them correct format for function
coords_extracted_unlisted <- coords_extracted

# make coords dataframe and label colnames
for(i in 1:length(coords_extracted_unlisted)){
  for(j in 1:length(coords_extracted_unlisted[[i]][["coords"]])){
      coords_extracted_unlisted[[i]][["coords"]][[j]] <- as.data.frame(coords_extracted_unlisted[[i]][["coords"]][[j]])
      colnames(coords_extracted_unlisted[[i]][["coords"]][[j]]) <- c("long","lat")
  }
}

## Reorder coordinates so that multiple polygon areas are processed last

# This is a non-essential step that helps with processing. It orders the coordinates so that those LADs with multiple coordinate polygons (e.g. those LADs that include many islands) are extracted last. This helps ensure that, in the case of an extraction failure, there is a high likelihood that most LADs have already been acquired. To elaborate, because multiple polygon areas take a long time to check and extract, it is preferable that they are extracted last, after the bulk of other data is already acquired. Not doing so would result in long processing times with little extraction process if there are failures.

# reorder coord list so that multiple polygon areas are last
coord_lengths <- matrix(nrow = length(coords_extracted_unlisted), ncol = 2)
# make list of lengths of each element in coord list
for(i in 1:length(coords_extracted_unlisted)){
  coord_lengths[i,1] <- i
  coord_lengths[i,2] <- length(coords_extracted_unlisted[[i]][["coords"]])
}
# order coord indexes based on coord element length
coord_lengths_ordered <- coord_lengths[order(coord_lengths[,2]),]
#coord_lengths_ordered

# order coord list based on ordered coord indexes
ordered_coords <- list()
for(i in 1:length(coords_extracted_unlisted)){
  ordered_coords[[i]] <- coords_extracted_unlisted[[coord_lengths_ordered[i]]]
  names(ordered_coords)[i] <- names(coords_extracted_unlisted)[coord_lengths_ordered[i]]
}

save(ordered_coords, file = "../data/ordered_coords.Rdata")

```

## Add missing force names for Northern Ireland, Scotland, and Northamptonshire

By checking 'missing_data_df' and sorting by 'Force_missing' column, it is possible to quickly identify that the areas that are missing data are Scotland, Northern Ireland, and Northamptonshire. Scotland and NI are missing because they have their own Polcie services. Northamptonshire is missing because of a mismatch of census codes with the newly defined LADs North and West Northamptonshire. The code below manually adds the missing data.

```{r}
for(i in 1:length(ordered_coords)){
  if(is.na(ordered_coords[[i]][["force"]])){
    if(ordered_coords[[i]][["country"]] == "Scotland"){
      ordered_coords[[i]][["force"]] <- "Police Scotland"
    }
    else if(ordered_coords[[i]][["country"]] == "Northern Ireland"){
      ordered_coords[[i]][["force"]] <- "Police Service of Northern Ireland"
    }
    else{
      ordered_coords[[i]][["force"]] <- "Northamptonshire"
    }
  }
}

# check that there are no more missing forces
count <- 0
for(i in 1:length(ordered_coords)){
  if(is.na(ordered_coords[[i]][["force"]])){
    count <- count + 1
  }
}
print(paste0(count, " missing forces remain"))
```

```{r}
# # check depth of coord lists 
# count <- 0
# for(i in 1:length(coords_extracted_unlisted)){
#   for(j in 1:length(coords_extracted_unlisted[[i]][["coords"]])){
#     if(is.list(coords_extracted_unlisted[[i]][["coords"]][[j]][[1]])){
#       count <- count + 1
#     }
#   }
# }
# count
# new_coords <- coords_extracted_unlisted
```


# Extraction function

This part of the script specifies and runs the function used to extract data from the Police API. It takes the formatted coordinates acquired above and queries the Police API via POST request for stop and search data for each LAD over a period of time specified by the user.

```{r}
# Acquire Stop and Search data for a list of Local Authority (LA) areas over a 
# time period. 

# Arguments

## coord_list: List of areas of interest. Each element of list must be a data 
## frame with column names "lat" and "long". A separate function is available
## to create this list.

## most_recent_month: Numeric value specifying most recent month (e.g., 8 for August)

## most_recent_year (YYYY): Numeric value specifying most recent year (e.g., 2021)

## If one or both of most_recent_month/year is not specified, the function will 
## query Police API for the most recent update and use this as a start point.

## num_months_backwards: Number of months backwards for which data are required. 

## wait_time: Number of seconds to wait between retries of unsuccessful requests.

## max_tries: Maximum number of retries following an unsuccessful request. Default 
## is 'Inf' because usually all data are desired and unsuccessful requests are most
## likely timeouts that can be resolved by retrying.

# Value

## Output is a list containing 'results' and 'missing_entries'. 'results' contains
## the queried data in a dataframe. 'missing_entries' contains a list of LAs and
## their corresponding index in coord_list for which no data records exist for
## the specified time period.

la_search_police_api <- function(coord_list, 
                                 most_recent_month = NULL, 
                                 most_recent_year = NULL, 
                                 num_months_backwards = NULL, 
                                 back_to_month = NULL,
                                 back_to_year = NULL,
                                 wait_time = 5, 
                                 max_tries = Inf){
  # initialise dataframes
  overall_output <- data.frame()
  no_entries_df <- data.frame(setNames(rep(list(NA), 2), c("Index","Name")))
  
  # h loop iterates over LAs
  for(h in 1:length(coord_list)){
    print(paste0("Started area ", h)) # report start (useful for debugging)
    
    la_name <- coord_list[[h]][["la_name"]] # LA name
    la_code <- coord_list[[h]][["la_code"]] # la census code
    
    county <- coord_list[[h]][["county"]] # county name
    if(is_empty(county)){
      county <- NA # set NA if missing
    }
    
    region <- coord_list[[h]][["region"]] # region name
    if(is_empty(region)){
      region <- NA # set NA if missing
    }
    
    country <- coord_list[[h]][["country"]] # country name
    
    force <- coord_list[[j]][["force"]]
    
    # get most recent update if most_recent data not specified
    if(is.null(most_recent_month) || is.null(most_recent_year)){ 
      # get most recent update from API:
      date <- httr::content(
        httr::GET("https://data.police.uk/api/crimes-street-dates"))[[1]][["date"]] 
      most_recent_month <- as.numeric(substr(date,6,7))
      most_recent_year <- as.numeric(substr(date,1,4))
    }
    else{
      most_recent_month <- most_recent_month
      most_recent_year <- most_recent_year
    }
    
    if(is.null(num_months_backwards)){
      num_months_backwards <- 1 + (12 * (most_recent_year - back_to_year)) + (most_recent_month - back_to_month)
    }
    
    area_output <- data.frame() # initialise area output df
    number_months_acquired <- 0 # initialise number of months acquired counter
    
    # i loop iterates over the months required 
    for(i in 1:num_months_backwards){
      month_output <- data.frame()
      # format date to what is needed for API query ("yyyy-mm")
      if(i == 1){ # set values for first iteration
        month_num <- most_recent_month
        year <- most_recent_year
      }
      else{ # subsequent iterations
        month_num <- month_num - 1 # backwards a month each iteration
        if(month_num %% 12 == 0){ # if reach a new year, start months from 12 again
          month_num <- 12
          year <- year - 1 # backwards a year
        }
      }
      if(month_num < 10){ # paste 0 for months lower than 10
        month <- paste("0", month_num, sep = "")
      }
      else{
        month <- month_num
      }
      
      date <- paste(year, "-", month, sep = "") # combine dates into one string
      
      # j loop iterates over the coordinate sets within each LA and creates a
      # polygon string to be searched, and then submits the query.
      # Most LAs have only one coordinate set, but some have multiple (e.g., 
      # those that include islands). The function therefore needs to search 
      # each coordinate set within a LA separately.
      coord_string <- c() # initialise vector for coord string
      for(j in 1:length(coord_list[[h]][["coords"]])){
        # set this iteration's coordinate set:
        area_coords <- coord_list[[h]][["coords"]][[j]]
        # combine coord strings into format required by API (much quicker than looping):
        coord_string <- paste0(area_coords$lat,",",area_coords$long, collapse = ":") 
        
        # create body for post request
        body <- list("poly" = coord_string,
                     "date" = date)
        # search API for this coordinate set and date:
        post_request <- httr::POST("https://data.police.uk/api/stops-street?", body = body)
        
        # if search quota reached, break (shouldn't be an issue but just in case)
        if(post_request[["status_code"]] == 429){
          print("Quota reached. Abandoning request.")
          break
        }
        else{
          # if the request didn't succeed, wait some time ('wait_time') and 
          # keep trying up until 'max_tries' attempts.
          attempt <- 1
          while(post_request[["status_code"]] != 200 && attempt <= max_tries){ 
            print(paste0("Server error. Trying again (", attempt,")"))
            Sys.sleep(wait_time) # wait some time before trying again
            try(
              post_request <- httr::POST("https://data.police.uk/api/stops-street?", body = body)
            )
            # if search quota reached, break (shouldn't be an issue but just in case)
            if(post_request[["status_code"]] == 429){ 
              print("Quota reached. Abandoning request.")
              break
            }
            attempt <- attempt + 1
          }
        }
        
        # get data from results of query
        df <- httr::content(post_request) 
        # unlist data and convert to dataframe:
        df_2 <- lapply(df, unlist)
        df_3 <- do.call(bind_rows, df_2)
        df_3$coord_set <- j # record which coordinate set data is from
        
        # add results of this coordinate set iteration (j) to the output for 
        # this month iteration (i). Use bindrows for (high) possibility that columns
        # from different iterations will be in different order/missing
        #area_output <- bind_rows(area_output, df_3) 
        month_output <- bind_rows(month_output, df_3)
        
        cat("\014") # clear console 
        # report overall (i.e., LA) progress
        print(paste0("Working... LA ", h, " of ", length(coord_list), 
                     " (", 
                     round(100 * (h / length(coord_list)), 2), "%)"))
        # report month progress
        print(paste0("Working... Month ", i, " of ", num_months_backwards, 
                     " (", date, ")"))
        # report coordinate set progress
        print(paste0("Working... ", j, " of ", 
                     length(coord_list[[h]][["coords"]]), 
                     " coordinate sets retrieved"))

        
      } # coordinate set loop (j) ends
    
      # if records have been acquired, increase months count
      if(nrow(month_output) > 0){
         number_months_acquired <- number_months_acquired + 1
      }
      
      # add data from this month to overall LA output
      area_output <- bind_rows(area_output, month_output)
      
    } # month loop (i) ends
    
    
    
    # If there were no records for this LA, record the LA iteration number
    # and name, then proceed to the next LA
    if(nrow(area_output) == 0){
      if(is.na(no_entries_df[1,1])){ # if first occurrence, replaces NAs 
        no_entries_df[1,] <- c(h, la_name)
      }
      else{ # rbind subsequent occurrences
        no_entries_df <- rbind(no_entries_df, c(h, la_name))        
      }
      print(paste0("No records for ", la_name))
      next # proceed to next LA
    }
    
    # add columns for LA name, county, region, country, and the iteration index
    # for the LA (useful for quickly identifying which LA the function reached 
    # if it breaks unexpectedly)
    area_output$la_name <- la_name
    area_output$la_code <- la_code
    area_output$county <- county
    area_output$region <- region
    area_output$country <- country
    area_output$index <- h
    area_output$number_months_acquired <- number_months_acquired
    area_output$proportion_months_acquired <- number_months_acquired / num_months_backwards
    area_output$force <- force
    
    # separate datetime into 2 columns
    area_output$time <- substr(area_output$datetime, 12, 19)
    area_output$date <- as.Date(substr(area_output$datetime, 1, 10))
    # set as time
    #area_output$time <- chron(times. =  df$time, format = c(times = "hh:mm:ss"))
    
    # move index and location data to front of df
    area_output <- area_output %>%
      select(index, la_name, la_code, county, region, country, everything()) %>%
      relocate(location.longitude, .before = location.latitude) %>% # put long next to lat
      relocate(date, .after = datetime) %>% 
      relocate(time, .after = date)

    overall_output <- bind_rows(overall_output, area_output)
    

    
    # create a temporary output list and save it every time a LA completes, so
    # that there is a backup in case function breaks. Saves to same folder as
    # script
    save_progress <- list(result = overall_output,
                          missing_entries = no_entries_df)
    save(save_progress, file = "./save_progress.Rdata")
  } # LA loop (h) ends
  
  # return output. 'result' is the data. 'missing_entries' provides a list of
  # LAs that are missing from the data because there were no records for the 
  # LA in the specified time period.
  return(list(result = overall_output,
              missing_entries = no_entries_df))
}
```


```{r}
# # for calculating nubmer of months to go back. now implemented in extraction function
# count_months <- function(start_month = NULL, start_year = NULL, end_month = NULL, end_year = NULL){
#   number_months_years <- 12 * (start_year - end_year)
#   number_months_months <- start_month - end_month
#   total_months <- number_months_years + number_months_months
#   return(total_months)
# }
# 
# count_months(start_month = 11, start_year = 2021, end_month = 12, end_year = 2018)
```

```{r}
# # for testing, select just one force (in this case Cleveland)
indexes <- c()
for(i in 1:length(ordered_coords)){
  if(ordered_coords[[i]][["force"]] == "Cleveland"){
    indexes <- append(indexes, i)
  }
}
just_cleveland <- ordered_coords[indexes]
```

## Run function

```{r}
# get rid of known missing entries for this period to reduce processing time next run
# load(file = "../data/stop_search_all_las_12_months_to_oct_21.Rdata")
# known_missing_entries <- data[[2]]
# ordered_coords_missing_removed <- within(ordered_coords, rm(list = known_missing_entries$Name))


test_coords <- ordered_coords[1:3]
data <- la_search_police_api(just_cleveland, back_to_month = 12, back_to_year = 2018)

df <- data[[1]] # extract the results data frame
df <- df[order(df$date,df$time),]
# save R object
#save(data, file = "../data/stop_search_all_las_12_months_to_oct_21.Rdata")
```


```{r}
# quick load of data collected from above function
# load(file = "../data/all_las_to_oct_21_incl_ni.Rdata")
# df <- data_new_coords[[1]]
```


